Inferring the Recombinant Genealogy of a Pandemic Pathogen in Real Time

### 2025-09-12 latex/main.tex
Add some methods for UShER conversion

% See https://github.com/jeromekelleher/sc2ts-paper/issues/628
% Referenced by the main Results.
The ARG captures highly prevalent deletions in SARS-CoV-2 (Table~\ref{tab:major_dels}).
Li et al.\cite{Li2023indels} performed a large-scale analysis of indels
using a global GISAID dataset (sampled up to August, 2022).
We examined the major indels reported in Table S1 of Li et al.\cite{Li2023indels}
which have a minimum frequency of 1\% (count of sequences divided by 9,149,680 sequences analyzed).
All of these indels are deletions and have high frequencies in the ARG
(sequence count divided by the total number of samples),
except a 9-base deletion at 11283, a 9-based deletion at 21987, and a 9-base deletion at 686.
During site remapping, the 9-base deletion at 11283 was split into
a 5-base deletion at 11283 (count of 339867) and a 4-base deletion at 11288 (count of 871782).

@@

\begin{table}
\centering
\caption{Major deletion events in the ARG.}
\label{tab:major_dels}
\begin{tabular}{rrcrcrr}
\toprule
Start & Length & Region & Node & Pango lineage & Count & Frequency (\%)\\
\midrule
28271 & 1 & non-coding & 1436808 & B.1 & 1118363 & 45.0561 \\
28248 & 6 & ORF8 & 220186 & B.1.617.2 (Delta) & 1114669 & 44.9073 \\
22029 & 6 & S / S1-NTD & 200039 & B.1.617.2 (Delta) & 1112410 & 44.8163 \\
11288 & 4 & ORF1ab / NSP6 & 1436802 & B.1.1.529 & 871782 & 35.1220 \\
28362 & 9 & N & 1436802 & B.1.1.529 & 871774 & 35.1216 \\
21633 & 9 & S / S1-NTD & 822854 & BA.2 (Omicron) & 532072 & 21.4359 \\
11292 & 5 & ORF1ab / NSP6 & 822854 & BA.2 (Omicron) & 531894 & 21.4287 \\
6513 & 3 & ORF1ab / NSP3 & 851246 & BA.1 (Omicron) & 340765 & 13.7286 \\
11283 & 5 & ORF1ab / NSP6 & 851246 & BA.1 (Omicron) & 339867 & 13.6924 \\
21765 & 6 & S / S1-NTD & 851246 & BA.1 (Omicron) & 339681 & 13.6849 \\
21988 & 8 & S / S1-NTD & 851246 & BA.1 (Omicron) & 339651 & 13.6837 \\
22194 & 3 & S / S1-NTD & 851246 & BA.1 (Omicron) & 312706 & 12.5982 \\
28271 & 1 & non-coding & 86456 & B.1.1.7 (Alpha) & 290446 & 11.7014 \\
11288 & 9 & ORF1ab / NSP6 & 86456 & B.1.1.7 (Alpha) & 290408 & 11.6998 \\
21991 & 3 & S / S1-NTD & 86456 & B.1.1.7 (Alpha) & 288831 & 11.6363 \\
21765 & 6 & S / S1-NTD & 86456 & B.1.1.7 (Alpha) & 290438 & 11.7010 \\
21765 & 6 & S / S1-NTD & 1265302 & BA.4 (Omicron) & 190425 & 7.6718 \\
28361 & 1 & N & 1246012 & BA.5.2.1 (Omicron) & 20822 & 0.8389 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}
\centering
\caption{Highly recurrent deletions.}
\label{tab:rec_dels}
\begin{tabular}{rrcr}
\toprule
Start & Length & Region & Occurrences \\
\midrule
29781 & 1 & non-coding & 12713 \\
29762 & 1 & non-coding & 11356 \\
29779 & 1 & non-coding & 10925 \\
29769 & 1 & non-coding & 10783 \\
29774 & 1 & non-coding & 10707 \\
29700 & 1 & non-coding & 8543 \\
29614 & 1 & ORF10 & 5041 \\
29555 & 1 & non-coding & 4816 \\
29543 & 1 & non-coding & 4806 \\
29541 & 1 & non-coding & 4804 \\
203 & 1 & non-coding & 3602 \\
222 & 1 & non-coding & 3441 \\
22013 & 6 & S / S1-NTD & 2111 \\
21991 & 3 & S / S1-NTD & 2087 \\
28254 & 1 & ORF8 & 1967 \\
27556 & 64 & ORF7a & 1281 \\
27579 & 3 & ORF7a & 1014 \\
28093 & 1 & ORF8 & 898 \\
21765 & 6 & S / S1-NTD & 478 \\
27532 & 1 & ORF7a & 403 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}
\centering
\caption{Most commonly observed deletions in the ARG or Li et al. (2023).} 
\label{tab:common_dels}
\begin{tabular}{rrcrrrr}
\toprule
\multicolumn{3}{c}{} & \multicolumn{2}{c}{Sc2ts ARG} & \multicolumn{2}{c}{Li et al. (2023)} \\
\cmidrule(lr){4-5} \cmidrule(lr){6-7}
Start & Length & Region & Count & Frequency (\%) & Count & Frequency (\%) \\
\midrule
28271 & 1 & non-coding & 1422469 & 57.3078 & 3853692 & 42.1183 \\
28248 & 6 & ORF8 & 1114712 & 44.9090 & 3026186 & 33.0742 \\
22029 & 6 & S / S1-NTD & 1114415 & 44.8970 & 3010665 & 32.9046 \\
11288 & 4 & ORF1ab / NSP6 & 871838 & 35.1242 & N/A & N/A \\
28362 & 9 & N & 871778 & 35.1218 & 3939164 & 43.0525 \\
21765 & 6 & S / S1-NTD & 825628 & 33.2625 & 3179887 & 34.7541 \\
21633 & 9 & S / S1-NTD & 532084 & 21.4364 & 1941962 & 21.2244 \\
11292 & 5 & ORF1ab / NSP6 & 531943 & 21.4307 & N/A & N/A \\
6513 & 3 & ORF1ab / NSP3 & 341714 & 13.7668 & 1984683 & 21.6913 \\
11283 & 5 & ORF1ab / NSP6 & 339935 & 13.6951 & 8360 & 0.0914 \\
21988 & 8 & S / S1-NTD & 339770 & 13.6885 & 590 & 0.0064 \\
22194 & 3 & S / S1-NTD & 313878 & 12.6454 & 1833050 & 20.0340 \\
11288 & 9 & ORF1ab / NSP6 & 311566 & 12.5522 & 3040576 & 33.2315 \\
21991 & 3 & S / S1-NTD & 308800 & 12.4408 & 912944 & 9.9779 \\
29762 & 1 & non-coding & 38030 & 1.5321 & N/A & N/A \\
29781 & 1 & non-coding & 36112 & 1.4549 & N/A & N/A \\
29769 & 1 & non-coding & 34315 & 1.3825 & N/A & N/A \\
29774 & 1 & non-coding & 33503 & 1.3498 & N/A & N/A \\
29779 & 1 & non-coding & 30515 & 1.2294 & N/A & N/A \\
11283 & 9 & ORF1ab / NSP6 & 60 & 0.0024 & 2011934 & 21.9891 \\
686 & 9 & ORF1ab / NSP1 & N/A & N/A & 115450 & 1.2618 \\
21987 & 9 & S / S1-NTD & N/A & N/A & 1918094 & 20.9635 \\
\bottomrule
\end{tabular}
\end{table}

>>>>

% See https://github.com/jeromekelleher/sc2ts-paper/issues/628
% Referenced by the main Results.
The ARG captures highly prevalent deletions in SARS-CoV-2 (Table~\ref{tab:major_dels}).

@@

Li et al.\cite{Li2023indels} performed a large-scale analysis of indels
using a global GISAID dataset (sampled up to August, 2022).
We examined the major indels reported in Table S1 of Li et al.\cite{Li2023indels}
which have a minimum frequency of 1\% (count of sequences divided by 9,149,680 sequences analyzed).
All of these indels are deletions and have high frequencies in the ARG
(sequence count divided by the total number of samples),
except a 9-base deletion at 11283, a 9-based deletion at 21987, and a 9-base deletion at 686.

\begin{table}
\centering
\caption{Major deletion events in the ARG.}
\label{tab:major_dels}
\begin{tabular}{rrcrcrr}
\toprule
Start & Length & Region & Node & Pango lineage & Count & Frequency (\%)\\
\midrule
28271 & 1 & non-coding & 1436808 & B.1 & 1118363 & 45.0561 \\
28248 & 6 & ORF8 & 220186 & B.1.617.2 (Delta) & 1114669 & 44.9073 \\
22029 & 6 & S / S1-NTD & 200039 & B.1.617.2 (Delta) & 1112504 & 44.8200 \\
11288 & 4 & ORF1ab / NSP6 & 1436802 & B.1.1.529 & 871782 & 35.1220 \\
28362 & 9 & N & 1436802 & B.1.1.529 & 871774 & 35.1216 \\
21633 & 9 & S / S1-NTD & 822854 & BA.2 (Omicron) & 532072 & 21.4359 \\
11292 & 5 & ORF1ab / NSP6 & 822854 & BA.2 (Omicron) & 531894 & 21.4287 \\
6513 & 3 & ORF1ab / NSP3 & 851246 & BA.1 (Omicron) & 340757 & 13.7283 \\
11283 & 5 & ORF1ab / NSP6 & 851246 & BA.1 (Omicron) & 339867 & 13.6924 \\
21765 & 6 & S / S1-NTD & 851246 & BA.1 (Omicron) & 339681 & 13.6849 \\
21987 & 9 & S / S1-NTD & 851246 & BA.1 (Omicron) & 339651 & 13.6837 \\
22194 & 3 & S / S1-NTD & 851246 & BA.1 (Omicron) & 312706 & 12.5982 \\
28271 & 1 & non-coding & 86456 & B.1.1.7 (Alpha) & 290446 & 11.7014 \\
11288 & 9 & ORF1ab / NSP6 & 86456 & B.1.1.7 (Alpha) & 290408 & 11.6998 \\
21991 & 3 & S / S1-NTD & 86456 & B.1.1.7 (Alpha) & 288831 & 11.6363 \\
21765 & 6 & S / S1-NTD & 86456 & B.1.1.7 (Alpha) & 290438 & 11.7010 \\
21765 & 6 & S / S1-NTD & 1265302 & BA.4 (Omicron) & 190425 & 7.6718 \\
686 & 9 & ORF1ab / NSP1 & 2698748 & BA.4 (Omicron) & 23330 & 0.9399 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}
\centering
\caption{Highly recurrent deletions.}
\label{tab:rec_dels}
\begin{tabular}{rrcr}
\toprule
Start & Length & Region & Occurrences \\
\midrule
29781 & 1 & non-coding & 12754 \\
29762 & 1 & non-coding & 11383 \\
29779 & 1 & non-coding & 10951 \\
29769 & 1 & non-coding & 10809 \\
29774 & 1 & non-coding & 10735 \\
29700 & 1 & non-coding & 8562 \\
686 & 9 & ORF1ab / NSP1 & 6436 \\
29614 & 1 & ORF10 & 5039 \\
29555 & 1 & non-coding & 4814 \\
29543 & 1 & non-coding & 4804 \\
29541 & 1 & non-coding & 4802 \\
203 & 1 & non-coding & 3604 \\
222 & 1 & non-coding & 3442 \\
21991 & 3 & S / S1-NTD & 2088 \\
28254 & 1 & ORF8 & 1969 \\
28093 & 1 & ORF8 & 900 \\
21765 & 6 & S / S1-NTD & 482 \\
27532 & 1 & ORF7a & 404 \\
335 & 1 & ORF1ab / NSP1 & 370 \\
337 & 1 & ORF1ab / NSP1 & 346 \\
\bottomrule
\end{tabular}
\end{table}
